---
title: "Lab 7: Functions + Fish"
author: "Tillman Erb"
date: "11/07/2024"
format: 
  html:  
    embed-resources: true
    code-tools: true
    toc: true
editor: source
code-fold: true
self-contained: true
theme: darkly
execute:
  echo: true
  error: true
  warning: false
---

The goal of this lab is learn more about exploring missing data and writing
modular code.

```{r}
#| label: setup

library(here)
library(tidyverse)

bf_fish <- read_csv(file = here::here("data","BlackfootFish.csv"))
glimpse(bf_fish)
```

## The Data

This lab's data concerns mark-recapture data on four species of trout from the
Blackfoot River outside of Helena, Montana. These four species are
**rainbow trout (RBT)**, **westslope cutthroat trout (WCT)**, **bull trout**,
and **brown trout**.

Mark-recapture is a common method used by ecologists to estimate a population's
size when it is impossible to conduct a census (count every animal). This method
works by *tagging* animals with a tracking device so that scientists can track
their movement and presence.

## Data Exploration

The measurements of each captured fish were taken by a biologist on a raft in
the river. The lack of a laboratory setting opens the door to the possibility of
measurement errors.

**1. Let's look for missing values in the dataset. Output ONE table that answers BOTH of the following questions:**

+ **How many observations have missing values?**
+ **What variable(s) have missing values present?**

::: callout-tip
# You should use `across()`!
:::

```{r}
#| label: find-missing-values
#| 
table_count <- bf_fish |>
  filter(if_any(.fns = is.na, .cols = everything())) |>
  summarize(count = n())

null_counts <- bf_fish |>
  filter(if_any(.cols = everything(), .fns = is.na)) |>
  summarize(across(.cols = everything(), .fns = ~ sum(is.na(.x)))) #Couldn't figure out how to apply the sum(is.na(.x)) to all columns without erroring. Learned about anonymous functions through https://gemini.google.com/app/43c895d2be4a0918

null_counts$total_rows_with_NAs <- table_count$count
null_counts

```

**2. Create ONE thoughtful visualization that explores the frequency of missing values across the different years, sections, and trips.**

```{r}
#| label: visual-of-missing-values-over-time
bf_fish |>
  filter(if_any(.cols = everything(), .fns = is.na)) |>
  mutate(trip = as.character(trip),
         trip = fct_recode(trip,
      "Trip 1" = "1",
      "Trip 2" = "2")) |>
  ggplot(mapping = aes(x = year, fill = section)) +
  facet_grid(~ trip) + 
  geom_bar() +
  scale_fill_viridis_d() + #Debugged here with Gemini #https://gemini.google.com/app/43c895d2be4a0918
  labs(x = "Year",
       y = NULL,
       title = "Missing observation frequency by section, trip and year",
       subtitle = "Number of missing observations",
       color = "Section")
```

## Rescaling the Data

If I wanted to rescale every quantitative variable in my dataset so that they
only have values between 0 and 1, I could use this formula:

</br>

$$y_{scaled} = \frac{y_i - min\{y_1, y_2,..., y_n\}}{max\{y_1, y_2,..., y_n\} 
- min\{y_1, y_2,..., y_n\}}$$

</br>

I might write the following `R` code to carry out the rescaling procedure for the `length` and `weight` columns of the `BlackfoorFish` data:

```{r}
#| echo: true
#| eval: false

fish <- fish |> 
  mutate(length = (length - min(length, na.rm = TRUE)) / 
           (max(length, na.rm = TRUE) - min(length, na.rm = TRUE)), 
         weight = (weight - min(weight, na.rm = TRUE)) / 
           (max(weight, na.rm = TRUE) - min(length, na.rm = TRUE)))
```

This process of duplicating an action multiple times can make it difficult to
understand the intent of the process. *Additionally, it can make it very difficult to spot mistakes.*

**3. What is the mistake I made in the above rescaling code?**

*When copy pasting the code (from length to weight), you forgot to change one of the 'length' values back to 'weight'*

When you find yourself copy-pasting lines of code, it's time to write a
function, instead!

**4. Transform the repeated process above into a `rescale_01()` function. Your function should...**

+ **... take a single vector as input.**
+ **... return the rescaled vector.**

```{r}
#| label: write-rescale-function

rescale_01 <- function(vec, na.rm = TRUE) {
  if (!is.numeric(vec[1])){
    stop("Error: vector provided is not numeric")
  } else if (length(vec) <= 1) {
    stop("Error: vector length must be greater than 1")
  }
  vec_range <- range(vec, na.rm = TRUE, finite = TRUE)
  vec <- (vec - vec_range[1]) / 
           (vec_range[2] - vec_range[1])
  if (na.rm){
    vec <- vec[!is.na(vec)] #https://www.geeksforgeeks.org/remove-na-values-from-vector-in-r/
  }
  return(vec)
}
```

::: callout-tip
# Efficiency 

Think about the efficiency of the function you wrote. Are you calling the
**same** function multiple times? You might want to look into the `range()` 
function. 
:::

**5. Let's incorporate some input validation into your function. Modify your previous code so that the function stops if ...**

+ **... the input vector is not numeric.**
+ **... the length of the input vector is not greater than 1.**

::: callout-tip
# Modify Previous Code

Do not create a new code chunk here -- simply add these stops to your function
above!
:::

## Test Your Function

**6. Run the code below to test your function. Verify that the maximum of your rescaled vector is 1 and the minimum is 0!**

```{r}
#| label: verify-rescale-function

x <- c(1:25, NA)

rescaled <- rescale_01(x)
min(rescaled, na.rm = TRUE)
max(rescaled, na.rm = TRUE)
```

Next, let's test the function on the `length` column of the `BlackfootFish` data.

**7. The code below makes a histogram of the original values of `length`. Add a plot of the rescaled values of `length`. Output your plots side-by-side, so the reader can confirm the only aspect that has changed is the scale.**

::: callout-warning
This will require you to call your `rescale_01()` function within a `mutate()`
statement in order to create a `length_scaled` variable.
:::

```{r}
#| label: compare-original-with-rescaled-lengths
#| layout-ncol: 2

bf_fish |>  
  ggplot(aes(x = length)) + 
  geom_histogram(binwidth = 45) +
  labs(x = "Original Values of Fish Length (mm)") +
  scale_y_continuous(limits = c(0,4000))

# Code for Q7 plot.

bf_fish |>
  mutate(length_scaled = rescale_01(length)) |>
  ggplot(aes(x = length_scaled)) + 
  geom_histogram(binwidth = 0.045) +
  labs(x = "Scaled Values of Fish Length (mm)") +
  scale_y_continuous(limits = c(0,4000)) 
```

::: callout-tip
1. Set the y-axis limits for both plots to go from 0 to 4000 to allow for direct comparison across plots.

2. Pay attention to `binwidth`!

3. Use a Quarto code chunk option to put the plots side-by-side.
:::

## Use Variables within a Dataset

Suppose you would like for your `rescale()` function to perform operations on a **variable within a dataset**. Ideally, your function would take in a data
frame and a variable name as inputs and return a data frame where the variable
has been rescaled.

**8. Create a `rescale_column()` function that accepts two arguments:**

+ **a dataframe**
+ **the name(s) of the variable(s) to be rescaled**

**The body of the function should call the original `rescale_01()` function you wrote previously. Your solution MUST use one of the `rlang` options from class.**

::: callout-tip
If you are struggling with this task, I recommend looking back over the 
[data frame functions](https://r4ds.hadley.nz/functions.html#data-frame-functions)
section of R for Data Science!
:::

```{r}
#| label: rescale-data-frame-function

rescale_column <- function(frm, col_vec) {
    frm |>
      mutate(across(.cols = {{col_vec}},
                    .fns = ~ rescale_01(.x, #Needed some help debugging here #https://gemini.google.com/app/d9e0a223f600c4a1
                                        na.rm = FALSE))
      )
}
```

**9. Use your `rescale_column()` function to rescale *both* the `length` and `weight` columns.**

::: callout-warning
I expect that you carry out this process by calling the `rescale_column()` function only ONE time!
:::

```{r}
#| label: rescale-two-columns
bf_fish |> 
  rescale_column(length:weight)
```
